---
title: "StationMapping"
author: "Patrick D. lorch"
date: "2022-09-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Rmotus package install

Probably do not need to run again.

```{r}
install.packages("motus", 
                 repos = c(birdscanada = 'https://birdscanada.r-universe.dev',
                           CRAN = 'https://cloud.r-project.org'))

# If individual package fail
install.packages("crayon")
install.packages("assertthat")
install.packages("cpp11")
install.packages("jsonlite")

# If this still fails you can do this and rerun
options(timeout = max(1000, getOption("timeout")))

```

## MotusRBook

Based on examples in **Motus R Book**
by Tara L. Crewe, Zoe Crysler, and Philip Taylor
https://beta.motus.org/MotusRBook/


```{r setup}
library(motus)
library(lubridate)
library(dplyr)
Sys.setenv(TZ = "UTC")

# If you need a project file, and an update run this.

proj.458.sql.motus <- tagme(projRecv = 458, new = TRUE, update = TRUE) # CA shorebirds
proj.247.sql.motus <- tagme(projRecv = 247, new = TRUE, update = TRUE) # SSRS

# If you need an updated project file, without activity, node data, or depricated batches data.

proj.458.sql.motus <- tagme(projRecv = 458, new = TRUE, update = TRUE) # CA shorebirds
proj.247.sql.motus <- tagme(projRecv = 247, 
                            new = FALSE, 
                            update = TRUE,
                            skipActivity = T,
                            skipNodes = T,
                            skipDeprecated = T) # SSRS

# If you just want to link to an existing project database and use the metadf.recvDeps from last download, use this.

proj.458.sql.motus <- tagme(projRecv = 458, 
                            dir = "C:/Users/PatrickLorch/OneDrive - SSRS/MotusGeneral/rmotus_work/Motus-work", 
                            new = FALSE, 
                            update = FALSE)
proj.247.sql.motus <- tagme(projRecv = 247, 
                            dir = "C:/Users/PatrickLorch/OneDrive - SSRS/MotusGeneral/rmotus_work/Motus-work", 
                            new = FALSE, 
                            update = FALSE)

# Receivers

# This will get data on all projects and receiver deployments
## Access this in recDeps table in proj.projnum.sql.motus
metadata(proj.458.sql.motus, projectIDs = 458)
metadata(proj.247.sql.motus, projectIDs = 247, delete = T)


motusLogout()
```

## Motus R book map example

I have struggled to get this example to work.  It requires lots of tweaking.
I am implementing a leaflet map instead below.

```{r data}
library(sf)
library(rnaturalearth)
library(ggplot2)

# get receiver metadata
tbl.458.recvDeps <- tbl(proj.458.sql.motus, "recvDeps")
df.458.recvDeps <- tbl.458.recvDeps %>% 
  collect() %>% 
  mutate(timeStart = as_datetime(tsStart),
         timeEnd = as_datetime(tsEnd),
         # for deployments with no end dates, make an end date a year from now
         timeEnd = if_else(is.na(timeEnd), Sys.time() + years(1), timeEnd))

tbl.247.recvDeps <- tbl(proj.247.sql.motus, "recvDeps")
df.247.recvDeps <- tbl.247.recvDeps %>% 
  collect() %>% 
  mutate(timeStart = as_datetime(tsStart),
         timeEnd = as_datetime(tsEnd),
         # for deployments with no end dates, make an end date a year from now
         timeEnd = if_else(is.na(timeEnd), Sys.time() + years(1), timeEnd))

# get running intervals for all receiver deployments
siteOp.458 <- with(df.458.recvDeps, interval(timeStart, timeEnd))
siteOp.247 <- with(df.247.recvDeps, interval(timeStart, timeEnd))

# set the date range you're interested in
dateRange <- interval(as_date("2014-01-01"), as_date("2017-12-31"))

# create new variable "active" which will be set to TRUE if the receiver was
# active at some point during your specified date range, and FALSE if not
df.458.recvDeps$active <- int_overlaps(siteOp.458, dateRange) 
df.247.recvDeps$active <- int_overlaps(siteOp.247, dateRange) 

# create map with receivers active during specified date range as red, and
# receivers with detections as yellow

# Run one of these to use in map
df.recvDeps = df.458.recvDeps
df.recvDeps = df.247.recvDeps
```

## Create Leaflet map with slider

From here: https://rdrr.io/cran/leaflet.extras2/man/addTimeslider.html

```{r leaflet}
# install.packages("leaflet.extras2", "geojsonsf")
library(leaflet)
library(leaflet.extras2)
library(sf)
library(geojsonsf)

# eliminate stations with no Lat, Long.
data = df.recvDeps[-which(is.na(df.recvDeps$latitude)),]
# eliminate station with no timeStart
data = data[-which(is.na(data$timeStart)),]

projcrs <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"

sf.recvDeps <- sf::st_as_sf(data,
                            coords = c("longitude", "latitude"),
                            crs = projcrs)
sf.recvDeps <- st_cast(sf.recvDeps, "POINT")
sf.recvDeps$time = sf.recvDeps$timeStart

# color palette
pal <- colorFactor(c("gray", "red", "white"), 
                   levels = c("terminated", "active", "pending"))

leaflet() %>%
  addTiles(group = "OSM (default)") %>%
  addProviderTiles('Esri.WorldImagery',group = "Esri World Imagery") %>%
  addTimeslider(data = sf.recvDeps,
                color = pal(sf.recvDeps$status),
                popup = ~paste0(name, ";  ", 
                                fixtureType, ";  ", 
                                lubridate::date(sf.recvDeps$timeStart)),
             options = timesliderOptions(
               position = "topright",
               timeAttribute = "time",
               alwaysShowDate = TRUE,
               range = TRUE)) %>%
  addLegend("bottomleft", pal = pal, values = sf.recvDeps$status,
                title = "Motus stations <br> (data: motus.org)",
                opacity = 0.5) %>%
      addLayersControl(
        baseGroups = c("OSM (default)", "Esri World Imagery"),
        position = "bottomright",
        options = layersControlOptions(collapsed = TRUE)) %>%
  setView(-119, 37, 5) # zoom into California

```

## Example using Leaflet and shiny

I found this first then found a leaflet.extras2 slider

This comes from here:
https://blogs.unimelb.edu.au/researcher-library/2020/06/19/making-a-covid-19-map-in-r-using-shiny-and-leaflet/

I never finished setting this up.

```{r rshinyleaflet}
# install and load necessary libraries
# install.packages(c(tidyverse", "shiny", "leaflet", "RColorBrewer"))
library(tidyverse)
library(shiny)
library(leaflet)
library(RColorBrewer)

# color palette
pal <- colorFactor(c("gray", "red", "white"), 
                   domain = c("terminated", "active", "pending"))

# specify shiny user interface and time slider
ui <- bootstrapPage(tags$style(type = "text/css", 
                               "html, body, .leaflet {width:100%; height:100%}"),
                    leafletOutput("map", width = "100%", height = "100%"),
                    # position and properties of the time slider
                    absolutePanel(bottom = 10, right = 300, draggable = TRUE,
                                  # slider title, step increments, and ticks
                                  sliderInput("integer", "Date range:",ticks = FALSE, 
                                              min = min(df.recvDeps$timeStart), 
                                              max = ~max(df.recvDeps$timeEnd), 
                                              value = 1:30, step = 1,
                                              animate = animationOptions(interval = 1000, 
                                                                         loop = TRUE))))

# shiny server input/output
server <- function(input, output, session) {
  filtereddf.recvDeps <- reactive({
    df.recvDeps %>%
      filter(day >= input$integer[1] & day <= input$integer[2])
  })
  output$map <- renderLeaflet({
    leaflet(df.recvDeps) %>%
      addTiles(group = "OSM (default)") %>%
      addProviderTiles('Esri.WorldImagery',group = "Esri World Imagery") %>%
      # set boundaries for map
      fitBounds(lng1 = min(df.recvDeps$longitude), 
                lat1 = min(df.recvDeps$latitude), 
                lng2 = max(df.recvDeps$longitude), 
                lat2 = max(df.recvDeps$latitude)) %>%
      # add legend for the map
      addLegend("bottomleft", pal = pal, values = ~status,
                title = "Motus stations <br> (data: motus.org)",
                opacity = 0.5,
                bins = 4)
  })
  observe({
    leafletProxy("map", df.recvDeps = filteredData()) %>%
      clearMarkers() %>%
      addCircleMarkers(lng = ~longitude,
                       lat = ~latitude,
                       # radius = ~log(value) * 3.5,
                       weight = 1,
                       opacity = 10,
                       color = pal(status),
                       popup = ~paste0(name; ",  ", fixtureType),
                       clusterOptions = markerClusterOptions()) %>%
      addLayersControl(
        baseGroups = c("OSM (default)", "Esri World Imagery"),
        options = layersControlOptions(collapsed = TRUE))

  }) 
}

```

